---
title: "U.S. Chronic Disease Indicators CDI"
author: "Team 3"
resource_files:
- .Renviron
output:
  html_notebook:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
runtime: shiny
---

```{r setup, include=FALSE}
library(tidyverse)
library(data.world)
library(shiny)
knitr::opts_chunk$set(echo = TRUE)
```

## **Github Link** 
[Team 3 github](https://github.com/KassPerez/DVProject2)

## **Introduction** 
The purpose of this project is to create the following data visualizations using R's ggplot2 library: Boxplot(s), Histograms(s), Scatter Plots(s), Parameter(s), and Barchart(s). 

## **Description of the dataset:** 
The U.S. Chronic Disease Indicators dataset contains over 30,000 entries. Each instance contains the attributes: yearstart, yearend, locationabbr, locationdesc, datasource, topic (disease type), question, response, datavalueunit, datavaluetype, datavalue, datavalue, datavaluealt, datavaluefootnotesymbol, datavaluefootnote, lowconfidencelimit, highconfidencelimit, strateficationcategory1, stratification1, stratificationcategory2, stratification2, stratificationcategory3, stratification3, geolocation, responseid, locationid, topicid, questionid, datavaluetypeid, strateficationcategoryid1, stratificationid1, stratificationcategoryid2, stratificationid2, stratificationcategoryid3, and stratificationid3.

```{r}
project <- "https://data.world/omarolivarez/us-chronic-disease-indicators"
data.world::set_config(cfg_env("DW_API"))
df <- data.world::query(data.world::qry_sql("SELECT * FROM u_s_chronic_disease_indicators_cdi LIMIT 30000"), dataset = project)
```

## One Green Thing with No Aggregate Measures 
### *Boxplot of Data Values* 
Below my team plotted boxplots of data values, which are counts of a disease per year. Additionally, we separated these by disease type, to search for large discrepancies between the occurrences of each disease type per year. Among all disease types, Chronic Kidney Disease stands out in that all of its data values are above all other data values for every other disease. Overarching Diseases stands out for the opposite reason: every one of its data values occur below the means for all other disease types. 

```{r}

renderPlot({
    ggplot(df) + geom_boxplot(mapping = aes(x=topic, y=datavaluealt)) +   theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
})
```

## Two Green Things with No Aggregate Measures 
### *Scatterplot of Diseases Across Confidence Ranges and Data Types*
Below our team plotted data_value against the confidence range, with color classified by the topic, or disease type. Confidence range actuallly does not exist in the original dataset, so our team wrote a SQL query to essentally create a calculated field composed of the difference between high_confidence_limit and low_confidence_limit. The scatter plot below demonstrates an interesting phenomena: data values occur exclusively above a positive slope line up until a confidence range of 40, where a few data values deviate slightly below the invisible slope.
```{r}
project <- "https://data.world/omarolivarez/us-chronic-disease-indicators"
data.world::set_config(cfg_env("DW_API"))
df2 <- data.world::query(data.world::qry_sql("SELECT yearstart, locationabbr as state, topic, datavaluealt as data_value,(highconfidencelimit - lowconfidencelimit) as confidence_range
FROM u_s_chronic_disease_indicators_cdi WHERE datavaluealt < 200 limit 30000"), dataset = project)
```

```{r}

renderPlot({
    ggplot(df2) + geom_point(mapping = aes(x=confidence_range, y=data_value, color=topic))
})
```

```{r}
project <- "https://data.world/omarolivarez/us-chronic-disease-indicators"
data.world::set_config(cfg_env("DW_API"))
df3 <- data.world::query(data.world::qry_sql("SELECT yearstart, datavaluealt as data_value
FROM u_s_chronic_disease_indicators_cdi GROUP BY yearstart limit 30000"), dataset = project)
```

## One Green Thing with Aggregate Measures 
### *Histogram of Frequency of Diseases*
Below we plot the frequency of disaeases, segmenting each bar by demographic information. It is apparent that the data collectors did a good job of ensuring that an even distribution of each of their stratification1 demographics were distributed in the data. It is also clear that Diabetes is several multiples higher than any of the other diseases.

```{r}

renderPlot({
  ggplot(df, mapping = aes(x = topic, fill = stratification1)) + geom_histogram(bins = 1, stat="count") +   theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
})

```
### *Histogram of Data values*
Here we see diseases across datavalues. Once again, it is very clear that Diabetes is the disease most frequently recorded, with that frequency increasing at data values greater than 50. We can see that the distribution of data values is not normally distributed, but instead has three large swells throughout the distribution. At the high and low end of data values (close to 0 and greater than 80), the frequency of data values drops noticeably.

```{r}

inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50, 100), selected = 20)
)

renderPlot({
  ggplot(df, mapping = aes(x = datavaluealt,fill = topic)) + geom_histogram(bins=input$n_breaks) + xlim(0, 100)
})

```



